Hybrid platforms and bargaining power

### 2024-01-12 src/paper/application.tex
Add extension for two-sided bargaining

@@

                \legend{$N_P$, $N_F$, $A$}
            \end{axis}
        \end{tikzpicture}
        \caption{Platform product variety and number of fringe entrants}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}
            \begin{axis}[
                    xlabel={$N_P$},
                    ylabel={},
                    xmin=0, xmax=1.5, % adjust these values as needed
                    ymin=0, ymax=1.2, % adjust these values as needed
                    axis lines=left,
                    xtick={0, 0.6},
                    legend pos=north west,
                    width=\linewidth,
                ]
                \addplot[color=myblue,mark=none,thick] table[x=N_P,y=pi_P]{\equilibriumFullSurplusTwoSided};
                \addplot[color=myred,mark=none,thick] table[x=N_P,y=pi_F]{\equilibriumFullSurplusTwoSided};
                \addplot[color=black,mark=none,thick,dashed] table[x=N_P,y=pi_P]{\equilibriumFullSurplusTwoSided};

                \legend{$\pi_P$, $\pi_F$, $CS$}
            \end{axis}
        \end{tikzpicture}
        \caption{Platform and fringe profits}
    \end{subfigure}
    \caption{Equilibrium outcomes in the case when the whole surplus is bargained over ($N_C = 0.6, V_P = 1, V_F = 1, I_F = 0.2$), and both consumers and fringe firms participate in the bargaining. As before, the platform's profits are increasing in $N_P$, while the fringe's profits are decreasing. On the other hand, consumer welfare is equal to platform profits, therefore it is increasing in $N_P$. The shaded region represents hybrid operation, while in the unshaded region, the platform operates in pure retail mode.}
    \label{fig:equilibrium_full_surplus_two_sided}
\end{figure}

>>>>

\subsection{Consumers also participate in bargaining}

In the second extension, I consider the case when the consumers also participate in the bargaining process.
This entails two changes compared to the main model.
First, the players bargain over the total surplus generated on all sides of the market, not just total profits.
Second, the outcomes are assumed to be described by a cooperative game with three types of players: the platform, the fringe firms, and the consumers.

Let us start by deriving total surplus as a function of $N_P, N_F$ and $N_C$.
Under the logit-like demand structure described in \cref{sec:example}, the total surplus is given by
\begin{proposition}
    \label{prop:profits_total_surplus}
    If the platform can extract consumer surplus, then the total surplus is given by
    \begin{align*}
        \Pi(N_P, N_F, N_C) = \mu N_C \left[ \frac{N_P V_P + N_F V_F}{N_P V_P + N_F V_F + 1} + \log(N_P V_P + N_F V_F + 1) \right].
    \end{align*}
\end{proposition}
\begin{proof}[Proof of \cref{prop:profits_total_surplus}]
    See \textcite{small1981applied} for a derivation for the discrete case.
    The continuous case is analogous.
\end{proof}
Note, that this expression is of the form $\Pi(N_C, N_P, N_F) = N_C f(N_P, N_F)$.
Furthermore, $f(N_P, N_F) = g(V_P N_P + V_F N_F)$ for an increasing, strictly concave $g$.

Next, let us consider the bargaining outcomes in this three-sided setting.
\Cref{sec:cooperative_game_two_sided} formally describes the corresponding cooperative game, and \cref{prop:profit_sharing_two_sided} establishes the resulting profit shares.
To summarize, the shares of the various players are given by
Then, the resulting profit shares are given by
\begin{align*}
    \pi_P(N_P, N_F, N_C) &= N_C \int_0^1 s f(N_P, s N_F) \ds, \\
    \pi_F(N_P, N_F, N_C) &= N_C \int_0^1 s^2 N_F \partial_2 f(N_P, s N_F) \ds \\
    CS(N_P, N_F, N_C) &= N_C \int_0^1 s f(N_P, s N_F) \ds.
\end{align*}

There are a number of things to note here.
First, and most importantly, the consumers' share from total surplus is equal to the platform's.\footnote{
    This is a consequence of the fact that the total surplus function is linear in the number of consumers.
}
As a consequence, what's good for the platform is also optimal for the consumers.
Therefore, letting the platform choose the number of its products maximizes consumer welfare, as well (but not necessarily total surplus).

Second, the expression for the fringe profits has the same structure as in \cref{ass:profit_sharing}.
Therefore, if we show that \cref{ass:single_crossing} holds, it in turn implies that all of the results in \cref{sec:results} apply here, as well.
The following lemma establishes this result.
\begin{lemma}
    \label{lem:profit_assumptions_total_surplus}
    $f(n_P, n_F)$ satisfies \cref{ass:single_crossing}.
    Furthermore,
    \begin{align*}
        \frac{\partial^2 f(n_P, n_F)}{\partial n_P \partial n_F} < 0.
    \end{align*}
\end{lemma}

With this proven, we can again rely on \cref{prop:aggregate_size_additive} to see that, in equilibrium, an increase in $N_P$ decreases $N_F$.
Furthermore, this decrease is so large that total product variety, and thus total surplus (which in turn is equal to total profits) also decreases (\cref{fig:fringe_entry_eq_full_surplus_two_sided}).\footnote{
    As long as the market is in the hybrid regime.
    If $N_F^* = 0$, then an increase in $N_P$ mechanically increases total aggregate, and thus total surplus.
}

Nevertheless, there is an important distinction compared to the previous results.
As before, platform profits are again increasing in $N_P$.
However, in this case, it also implies that consumers' share is also increasing in this case (\cref{fig:profits_full_surplus_two_sided}).
Therefore, consumers do prefer hybrid operation to the the platform being a pure marketplace.

                \legend{$N_P$, $N_F$, $A$}
            \end{axis}
        \end{tikzpicture}
        \caption{Platform product variety and number of fringe entrants}
        \label{fig:fringe_entry_eq_full_surplus_two_sided}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}
            \begin{axis}[
                    xlabel={$N_P$},
                    ylabel={},
                    xmin=0, xmax=1.5, % adjust these values as needed
                    ymin=0, ymax=1.2, % adjust these values as needed
                    axis lines=left,
                    xtick={0, 0.6},
                    legend pos=north west,
                    width=\linewidth,
                ]
                \addplot[color=myblue,mark=none,thick] table[x=N_P,y=pi_P]{\equilibriumFullSurplusTwoSided};
                \addplot[color=myred,mark=none,thick] table[x=N_P,y=pi_F]{\equilibriumFullSurplusTwoSided};
                \addplot[color=black,mark=none,thick,dashed] table[x=N_P,y=pi_P]{\equilibriumFullSurplusTwoSided};

                \legend{$\pi_P$, $\pi_F$, $CS$}
            \end{axis}
        \end{tikzpicture}
        \caption{Platform and fringe profits}
        \label{fig:profits_full_surplus_two_sided}
    \end{subfigure}
    \caption{Equilibrium outcomes in the case when the whole surplus is bargained over ($N_C = 0.6, V_P = 1, V_F = 1, I_F = 0.2$), and both consumers and fringe firms participate in the bargaining. As before, the platform's profits are increasing in $N_P$, while the fringe's profits are decreasing. On the other hand, consumer welfare is equal to platform profits, therefore it is increasing in $N_P$. The shaded region represents hybrid operation, while in the unshaded region, the platform operates in pure retail mode.}
    \label{fig:equilibrium_full_surplus_two_sided}
\end{figure}

@@
